### 2021-04-28 submission/manuscript.Rmd
Figure legend edits and additions

**5-day laxative treatment leads to prolonged *C. difficile* colonization in mice.** We compared mice treated with the osmotic laxative PEG 3350 to our standard 10 mg/kg clindamycin treatment, which temporarily renders the mice susceptible to *C. difficile*, with mice typically clearing *C. difficile* within 10 days post-infection [@Tomkovich2019; @Tomkovich2020]. All PEG-treated mice were administered a 15% PEG solution in the drinking water for 5-days, one group was also treated with clindamycin, and one group was allowed to recover for 10 days prior to challenge (Fig. 1A). PEG treatment resulted in weight loss in all 3 groups of PEG-treated mice, with the greatest change in weight observed on the fifth day of PEG treatment and the mice recovered most of the weight five days after treatment (Fig. 1B). After either PEG, clindamycin, or PEG and clindamycin treatment all mice were challenged with 10^3^ *C. difficile* 630 spores (Fig. 1A). All treatments rendered mice susceptible to *C. difficile* colonization. However, PEG-treated mice remained colonized with *C. difficile* at a high level through thirty days post-infection (Fig. 1C). In contrast, the clindamycin-treated mice cleared *C. difficile* within ten days post-infection. Surprisingly, PEG-treated mice were still susceptible to *C. difficile* infection after 10-days of recovery from treatment, although *C. difficile* was not detectable in most of the group in the initial five days post-infection (Fig. 1C). From 8 days post-infection onward, the median *C. difficile* CFU stabilized for the 10-day recovery group of PEG-treated mice and remained high through 30 days post-infection (Fig. 1C). Thus, osmotic laxative treatment alone was sufficient to render mice susceptible to prolonged *C. difficile* colonization and PEG-treated mice remained susceptible up to ten days post-treatment.

@@

```
**5-day laxative treatment does not promote more severe CDIs despite altering the mucosal microbiota.** Given the findings from a previous study that demonstrated PEG treatment disrupts the mucus layer and alters the immune response in mice [@Tropini2018], we decided to examine the impact of PEG treatment on the mucosal microbiota and CDI severity. To evaluate the mucosal microbiota, we sequenced snips of tissue collected from the cecum, proximal colon, and distal colon. Similar to what was observed with the stool samples, alpha diversity was lower in the PEG-treated mice compared to clindamycin treatment (Fig. 3A). Although alpha diversity continued to increase over time based on the communities from PEG-treated mice collected at 20 and 30 days post-infection (Fig. 3A, Data Set S1, Sheet X). Group, time point, and their interactions with other variables (cage, experiment number, and sample type) explained the majority of the variation observed in mucosal communities (PERMANOVA combined R^2^ = `r f_d_top_R2_sum_t`, *P* `r format.pval(f_d_top_p_t, eps = 0.05)`, Fig. 3B, Data Set S1, Sheet X). We saw the greatest difference in the mucosal microbiota between treatment groups at 6 days post-infection with 16 genera that were significantly different in all three of the tissue types we collected (cecum, proximal colon, and distal colon; Fig. S3A, Data Set S1, Sheet X). However, at 30 days post-infection only *Bacteroides*, *Clostridiales*, *Firmicutes*, and *Ruminococcaceae* were different between treatment groups and only in the cecum tissues (Fig. 3C, Fig. 2E, Data Set S1, Sheet X). Interestingly, one of the bacteria that was different between treatment groups at 6 days post-infection was *Peptostreptococcaceae* (the genera with an OTU sequence that matches *C. difficile*). *Peptostreptococcaceae* was primarily only present in the 5-day PEG treatment group of mice and decreased in the proximal and distal colon tissues over time (Fig. S2B). Thus, PEG treatment had a significant impact on the mucosal microbiota and we detected *C. difficile* sequences in the cecum, proximal colon, and distal colon tissue communities.

@@

Previous work examining the impact of PEG on the murine microbiota found that PEG treatment resulted in the permanent loss of S24-7, also known as *Muribaculum intestinale* [@Tropini2018]. We decided to check our *Porphyromonadaceae* OTUs because *Muribaculum intestinale* is classified as *Porphyromonadaceae* by the Ribosomal Database Project (RDP) database [@Vornhagen2020] and *Porphyromonadaceae* was a top feature in the random forest model predicting prolonged *C. difficile* colonization. We identified 4 OTUs that had at least 92% identity to *Muribaculum intestinale* and examined their abundance in mice that either cleared or were still colonized with *C. difficile* at 10 days post-infection. While all of the OTUs, were decreased by PEG and clindamycin treatment, there was some recovery in the mice that cleared (Fig. S4A). We also examined other *Porphyromonadaceae* and *Lachnospiraceae* OTUs since these were the 2 genera that were important to our classification model and contained multiple OTUs that were different at 5 days post-infection between mice that either cleared or remained colonized with *C. difficile* by 10 days post-infection (Data Set S1, sheet X). While individual *Porphyromonadaceae* and *Lachnospiraceae* OTUs tended to be more abundant in the mice that clear *C. difficile* relative to the mice that exhibit prolonged colonization (Fig. S4B-C), there is no single OTU that fits the pattern we observed at the genus level (Fig. 7C), suggesting multiple *Porphyromonadaceae* and *Lachnospiraceae* OTUs influenced *C. difficile* clearance. Overall, our results suggest that specific bacterial community differences explain the prolonged *C. difficile* colonization we observed in 5-day PEG and post-CDI 1-day PEG treated mice.

@@

@@

**Figure 1. 5-day PEG treatment prolongs susceptibility and mice become persistently colonized with *C. difficile*.** A. Setup of the experimental timeline for experiments with 5-day PEG treated mice. Clindamycin was administered at 10 mg/kg by intraperitoneal injection. 15% PEG 3350 was administered in the drinking water for five days. B. Weight change from baseline weight in groups after treatment with PEG and/or clindamycin, followed by *C. difficile* challenge. C. *C. difficile* CFU/gram stool measured over time (N = `r five_d_cfu_n_min`-`r five_d_cfu_n_max` mice per time point) via serial dilutions. The black line represents the limit of detection for the first serial dilution. CFU quantification data was not available for each mouse due to stool sampling difficulties (particularly the day the mice came off of the PEG treatment) or early deaths. For B-C, lines represent the median for each treatment group and circles represent samples from individual mice. Asterisks indicate time points where the weight change or CFU/g was significantly different between groups by the Kruskal-Wallis test with Benjamini-Hochberg correction for testing multiple time points. The data presented are from a total of `r five_day_total_exp` separate experiments.
\newpage

\includegraphics{figure_2.pdf}
**Figure 2. 5-day PEG treatment disrupts the stool microbiota for a longer amount of time compared to clindamycin-treated mice.** A. PCoA of Bray-Curtis distances from stool samples collected throughout the experiment.
\newpage

\includegraphics{figure_3.pdf}
**Figure 3. 5-day PEG treatment does not result in more severe CDIs, although mucosal microbiota is altered.** A.
\newpage

@@

```
\includegraphics{figure_4.pdf}
**Figure 4. 1-day PEG treatment renders mice susceptible to transient *C. difficile* colonization.** A. Setup of the experimental timeline for the 1-day PEG treated subset of mice. B. CFU/gram stool measured over time (N = `r one_day_sample_min`-`r one_day_sample_max` mice per time point) via several dilutions. The black dotted line represents the limit of detection for the first serial dilution. Asterisks indicate time points where the CFU/gram was significantly different between groups using the Kruskall-Wallis test with a Benjamini-Hochberg correction for multiple time points. C. Principle Coordinate Analysis plot of the groups over time with the alpha representating the same time scale as in panel D (`r one_day_top_contrib_name`: R^2^ = `r top_R2`; `r one_day_sec_contrib_name`: R^2^ = `r sec_R2`). D. Shannon diversity Index of the groups over time.  Only days with samples from all groups are shown. E. Line plots of relative percent abundance of selected genera over time. Only days with samples from all groups shown. The gray line represents the limit of detection.

``` {r figure_5_permanova}
fig_5_permanova_results = read_tsv(here("data/process/post_CDI_PEG_permanova_stools.tsv"))
r_sq_ordered_perm = fig_5_permanova_results %>% arrange(desc(R2)) %>%
  select(`effects`, `R2`, `Pr(>F)`) %>%
  filter(`Pr(>F)` < .05) #filter for sig
fig_5_permanova_sig_P_val <- r_sq_ordered_perm[,3] %>% sum()
fig_5_permanova_sig_R2 <- r_sq_ordered_perm[,2] %>% sum() #0.9280951
```

\includegraphics{figure_5.pdf}
**Figure 5. 1-day PEG treatment post *C. difficile* challenge prolongs colonization regardless of whether an FMT is also administered.** All mice were administered 10 mg/kg clindamycin intraperitoneally (IP) 1 day before challenge with *C. difficile* 630 spores on day 0. 15% PEG 3350 was administered in the drinking water for 1 day. Mice feces were collected daily through the end of the experiment (30 days post-infection for Clind. + 1-day PEG group, 15 days post-infection for all others). CFU quantification data was not available for each mouse due to stool sampling difficulties (particularly the day the mice came off of the PEG treatment) or early deaths. B. CFU/g of *C. difficile* stool measured over time via serial dilutions. C. Median and individual Shannon Diversity measured over time. D. Median and individual richness measured over time. B-D. Asterisks indicate time points with significant differences between groups using Kruskall-Wallis tests with a Benjamini-Hochberg correction for testing multiple times points. Black line represents the limit of detection for the first serial dilution. Background shading indicates PEG treatment period for applicable groups. Opacity of points reflects time point. The data presented are from a total of *** separate experiments.
\newpage

\includegraphics{figure_6.pdf}
**Figure 6. For 1-day PEG treatment post *C. difficile* challenge mice that also receive an FMT only some bacterial genera were restored.** A. PCoA of all groups + FMT plotted over time. Opacity of points reflects time point. B. Relative abundance of genera significantly different over multiple time points post FMT or PBS treatment, plotted over time. Limit of detection is .1%. C. Relative abundances of top 6 significant genera ranked by number of days significant, plotted over time. Limit of detection is .1%.

@@

\includegraphics{figure_7.pdf}
**Figure 7. Specific microbiota features associated with prolonged *C. difficile* colonization in PEG treated mice.** A.
\newpage

\includegraphics{figure_8.pdf}
**Figure 8. Schematic summarizing findings.** A.
\newpage

\includegraphics{figure_S1.pdf}
**Figure S1. 5-day PEG treatment plus 10-day recovery mice microbiota dynamics post-infection.** A.
\newpage

\includegraphics{figure_S2.pdf}
**Figure S2. Impact of PEG treatment on the mucosal microbiota is greatest 6 days post-infection**  
\newpage

@@

\includegraphics{figure_S4.pdf}
**Figure S4. Specific OTUs associated with clearance that are mostly absent in mice with prolonged *C. difficile* colonization. Ex. *Muribaculum intestinale*.** A.
\newpage

>>>>

**5-day laxative treatment leads to prolonged *C. difficile* colonization in mice.** We compared mice treated with the osmotic laxative PEG 3350 to our standard 10 mg/kg clindamycin treatment, which temporarily renders the mice susceptible to *C. difficile*, with mice typically clearing *C. difficile* within 10 days post-infection [@Tomkovich2019; @Tomkovich2020]. All PEG-treated mice were administered a 15% PEG solution in the drinking water for 5-days, one group was also treated with clindamycin, and one group was allowed to recover for 10 days prior to challenge (Fig. 1A). PEG treatment resulted in weight loss in all 3 groups of PEG-treated mice, with the greatest change in weight observed on the fifth day of PEG treatment and the mice recovered most of the weight five days after treatment (Fig. 1B). After either PEG, clindamycin, or PEG and clindamycin treatment all mice were challenged with 10^5^ *C. difficile* 630 spores (Fig. 1A). All treatments rendered mice susceptible to *C. difficile* colonization. However, PEG-treated mice remained colonized with *C. difficile* at a high level through thirty days post-infection (Fig. 1C). In contrast, the clindamycin-treated mice cleared *C. difficile* within ten days post-infection. Surprisingly, PEG-treated mice were still susceptible to *C. difficile* infection after 10-days of recovery from treatment, although *C. difficile* was not detectable in most of the group in the initial five days post-infection (Fig. 1C). From 8 days post-infection onward, the median *C. difficile* CFU stabilized for the 10-day recovery group of PEG-treated mice and remained high through 30 days post-infection (Fig. 1C). Thus, osmotic laxative treatment alone was sufficient to render mice susceptible to prolonged *C. difficile* colonization and PEG-treated mice remained susceptible up to ten days post-treatment.

@@

```
**5-day laxative treatment does not promote more severe CDIs despite altering the mucosal microbiota.** Given the findings from a previous study that demonstrated PEG treatment disrupts the mucus layer and alters the immune response in mice [@Tropini2018], we decided to examine the impact of PEG treatment on the mucosal microbiota and CDI severity. To evaluate the mucosal microbiota, we sequenced snips of tissue collected from the cecum, proximal colon, and distal colon. Similar to what was observed with the stool samples, alpha diversity was lower in the PEG-treated mice compared to clindamycin treatment (Fig. 3A). Although alpha diversity continued to increase over time based on the communities from PEG-treated mice collected at 20 and 30 days post-infection (Fig. 3A, Data Set S1, Sheet X). Group, time point, and their interactions with other variables (cage, experiment number, and sample type) explained the majority of the variation observed in mucosal communities (PERMANOVA combined R^2^ = `r f_d_top_R2_sum_t`, *P* `r format.pval(f_d_top_p_t, eps = 0.05)`, Fig. 3B, Data Set S1, Sheet X). We saw the greatest difference in the mucosal microbiota between treatment groups at 6 days post-infection with 16 genera that were significantly different in all three of the tissue types we collected (cecum, proximal colon, and distal colon; Fig. S3A, Data Set S1, Sheet X). However, at 30 days post-infection only *Bacteroides*, *Clostridiales*, *Firmicutes*, and *Ruminococcaceae* were different between treatment groups and only in the cecum tissues (Fig. 3C, Fig. 2E, Data Set S1, Sheet X). Interestingly, one of the bacteria that was different between treatment groups at 6 days post-infection was *Peptostreptococcaceae* (the genera that contains a sequence that matches *C. difficile*). *Peptostreptococcaceae* was primarily only present in the 5-day PEG treatment group of mice and decreased in the proximal and distal colon tissues over time (Fig. S2B). Thus, PEG treatment had a significant impact on the mucosal microbiota and we detected *C. difficile* sequences in the cecum, proximal colon, and distal colon tissue communities.

@@

Previous work examining the impact of PEG on the murine microbiota found that PEG treatment resulted in the permanent loss of S24-7, also known as *Muribaculum intestinale* [@Tropini2018]. We decided to check our *Porphyromonadaceae* operational taxonomic units (OTUs) because *Muribaculum intestinale* is classified as *Porphyromonadaceae* by the Ribosomal Database Project (RDP) database [@Vornhagen2020] and *Porphyromonadaceae* was a top feature in the random forest model predicting prolonged *C. difficile* colonization. We identified 4 OTUs that had at least 92% identity to *Muribaculum intestinale* and examined their abundance in mice that either cleared or were still colonized with *C. difficile* at 10 days post-infection. While all of the OTUs, were decreased by PEG and clindamycin treatment, there was some recovery in the mice that cleared (Fig. S4A). We also examined other *Porphyromonadaceae* and *Lachnospiraceae* OTUs since these were the 2 genera that were important to our classification model and contained multiple OTUs that were different at 5 days post-infection between mice that either cleared or remained colonized with *C. difficile* by 10 days post-infection (Data Set S1, sheet X). While individual *Porphyromonadaceae* and *Lachnospiraceae* OTUs tended to be more abundant in the mice that clear *C. difficile* relative to the mice that exhibit prolonged colonization (Fig. S4B-C), there is no single OTU that fits the pattern we observed at the genus level (Fig. 7C), suggesting multiple *Porphyromonadaceae* and *Lachnospiraceae* OTUs influenced *C. difficile* clearance. Overall, our results suggest that specific bacterial community differences explain the prolonged *C. difficile* colonization we observed in 5-day PEG and post-CDI 1-day PEG treated mice.

@@

#Figure out which experiments eneded at 4, 6, or 30 days post-infection
five_day_mice_samples <- five_day_mice %>% 
  group_by(exp_num, day, sample_type) %>% 
  tally()
#Figure out number of stool samples from 4,6, 10, 20 and 30 days post-infection
five_day_mice_samples <- five_day_mice %>% 
  filter(sample_type == "stool" & day %in% c(4, 6, 10, 20, 30)) %>% 
  group_by(day) %>% 
  tally()

@@

**Figure 1. 5-day PEG treatment prolongs susceptibility and mice become persistently colonized with *C. difficile*.** A. Setup of the experimental time line for experiments with 5-day PEG treated mice consisting of 4 treatment groups. 1. Clindamycin was administered at 10 mg/kg by intraperitoneal injection. 2. 15% PEG 3350 was administered in the drinking water for five days. 3. 5-day PEG plus clindamycin treatment. 4. 5-day PEG plus 10-day recovery treatment. All treatment groups were then challenged with 10^5^ *C. difficile* 630 spores. A subset of mice were euthanized on either 4 or 6 days post-infection and tissues were collected for histopathology analysis, the remaining mice were followed through 20 or 30 days post-infection. B. Weight change from baseline weight in groups after treatment with PEG and/or clindamycin, followed by *C. difficile* challenge. C. *C. difficile* CFU/gram stool measured over time (N = `r five_d_cfu_n_min`-`r five_d_cfu_n_max` mice per time point) via serial dilutions. The black line represents the limit of detection for the first serial dilution. CFU quantification data was not available for each mouse due to stool sampling difficulties (particularly the day the mice came off of the PEG treatment) or early deaths. For B-C, lines represent the median for each treatment group and circles represent samples from individual mice. Asterisks indicate time points where the weight change or CFU/g was significantly different between groups by the Kruskal-Wallis test with Benjamini-Hochberg correction for testing multiple time points. The data presented are from a total of `r five_day_total_exp` separate experiments.
\newpage

\includegraphics{figure_2.pdf}
**Figure 2. 5-day PEG treatment disrupts the stool microbiota for a longer amount of time compared to clindamycin-treated mice.** A. Principal Coordinate analysis (PCoA) of Bray-Curtis distances from stool samples collected throughout the experiment. For A and C, each circle represents a sample from an individual mouse and the transparency of the symbol corresponds to the day post-infection. B. Bray-Curtis distances of stool samples collected on either day 10 or 30 post-infection relative to the baseline sample collected for each mouse (before any drug treatments were administered). C. Shannon diversity in stool communities over time. The line indicates the median value for each treatment group. The colors of the symbols and lines represent the four treatment groups. D. 14 of the `r f_d_group_total_genera` genera affected by PEG treatment  (Data Set S1, sheet X). The symbols represent the median relative abundance for a treatment group at either baseline (open circle) or 1-day post treatment (closed circle). Data from the 5-day PEG and 5-day PEG plus 10-day recovery groups were analyzed by paired Wilcoxan signed-rank test with Benjamini-Hochberg correction for testing all identified genera. The clindamycin and 5-day PEG plus clindamycin treatment groups are shown for comparison. E. 6 of the `r f_d_group_multi_day` genera that were significantly different between the four treatment groups over multiple time points. Differences between treatment groups were identified by Kruskal-Wallis test with Benjamini-Hochberg correction for testing all identified genera. The gray vertical line (D) and horizontal vertical lines (E) indicate the limit of detection.
\newpage

\includegraphics{figure_3.pdf}
**Figure 3. 5-day PEG treatment does not result in more severe CDIs, although mucosal microbiota is altered.** A. Shannon diversity in cecum communities over time. The line indicates the median value for each treatment group. The colors of the symbols and lines represent the four treatment groups. A similar pattern was observed with the proximal and distal colon communities (Data Set S1, sheet X-X). B. PCoA of Bray-Curtis distances from mucosal samples collected throughout the experiment. Circles, triangles, and squares indicate cecum, proximal colon, and distal colon communities, respectively. For A-B, transparency of the symbol corresponds to the day post-infection that the sample was collected. C. The median relative abundance of the 4 genera that were significantly different between the cecum communities of different treatment groups on day 6 and day 30  (Data Set S1, sheet X). The gray vertical line indicate the limit of detection. D-E. The histopathology summary scores from cecum and colon H&E stained slides. The summary score is the total score based on evaluation of edema, cellular infiltration, and inflammation. Each category is given a score ranging from 0-4, thus the maximum possible summary score is 12. The tissue for histology was collected at either 4 (D) or 6 (E) days post-infection with the exception that one set of 5-day PEG treated mock-challenged mice were collected on day 0 post-infection (first set of open circles in D). Histology data were analyzed with the Kruskal-Wallis test followed by pairwise Wilcoxon comparisons with Benjamini-Hochberg correction.
\newpage

@@

```
\includegraphics{figure_4.pdf}
**Figure 4. 1-day PEG treatment renders mice susceptible to transient *C. difficile* colonization.** A. Setup of the experimental time line for the 1-day PEG treated  mice consisting of 3 treatment groups. 1. Clindamycin was administered at 10 mg/kg by intraperitoneal injection. 2. 15% PEG 3350 was administered in the drinking water for 1 day. 3. 1-day PEG plus 1-day recovery. The three treatment groups were then challenged with 10^5^ *C. difficile* 630 spores. B. *C. difficile* CFU/gram stool measured over time (N = `r one_day_sample_min`-`r one_day_sample_max` mice per time point) via several dilutions. The black dotted line represents the limit of detection for the first serial dilution. Asterisks indicate time points where the CFU/gram was significantly different between treatment groups by Kruskall-Wallis test with Benjamini-Hochberg correction for testing multiple time points. C. PCoA of Bray-Curtis distances from stool communities collected from the three treatment groups over time (`r one_day_top_contrib_name`: R^2^ = `r top_R2`; `r one_day_sec_contrib_name`: R^2^ = `r sec_R2`). D. Shannon diversity in stool communities over time with colored lines representing the median value for each treatment group. For B-D, each symbol represents a sample from an individual mouse and symbol transparency corresponds to the day post-infection that the sample was collected. E. Median relative abundances per treatment group for 6 out of the `r recovered_genera` genera that were affected by treatment, but recovered close to baseline levels by 7 days post-infection (Fig. 3E, Data Set S1, Sheet X). Stool samples from either baseline and day 1 or baseline and day 7 were analyzed by paired Wilcoxan signed-rank test with Benjamini-Hochberg correction for testing all identified genera. The gray horizontal line represents the limit of detection.

``` {r figure_5_permanova}
fig_5_permanova_results = read_tsv(here("data/process/post_CDI_PEG_permanova_stools.tsv"))
r_sq_ordered_perm = fig_5_permanova_results %>% arrange(desc(R2)) %>%
  select(`effects`, `R2`, `Pr(>F)`) %>%
  filter(`Pr(>F)` < .05) #filter for sig
fig_5_permanova_sig_P_val <- r_sq_ordered_perm[,3] %>% sum()
fig_5_permanova_sig_R2 <- r_sq_ordered_perm[,2] %>% sum() #0.9280951

#Number of experiments represented in [post-CDI PEG subset
#Use same code from post-CDI_PEG_subset() in code/utilities.R to determine number of samples we had C. difficile CFU counts for per day
post_CDI_mice <- metadata %>%     
  filter(group == "C" & exp_num %in% c("M7","M9")| #Only use C mice from these experiments. Allocated groups to figures based on paper outline.
             group == "CWM" & exp_num %in% c("M6","M7", "M9")|
             group == "RM" & exp_num %in% c("M7","M9")|
             group == "FRM" & exp_num %in% c("M9"))
#Total number of experiments in post-CDI subset
post_CDI_total_exp <- post_CDI_mice %>%
  filter(!duplicated(exp_num)) %>% #Remove duplicate experiment numbers
  tally() #Total number of unique experiments
```

\includegraphics{figure_5.pdf}
**Figure 5. 1-day PEG treatment post *C. difficile* challenge prolongs colonization regardless of whether an FMT is also administered.** A. Setup off the experimental time line for experiments with post-CDI PEG treated mice. There were a total of 4 different treatment groups. All mice were administered 10 mg/kg clindamycin intraperitoneally (IP) 1 day before challenge with 10^3-5^ *C. difficile* 630 spores. 1. Did not receive any additional treatment. 2. Immediately after *C. difficile* challenge, mice received 15% PEG 3350 in the drinking water for 1 day. 3-4. 3-days after challenge, mice received 1-day PEG treatment and then received either a fecal microbiota transplant (3) or PBS (4) solution by oral gavage.  Mice were followed through 15-30 days post-infection (only the post-CDI 1-day PEG group was followed through 30 days post-infection). B. CFU/g of *C. difficile* stool measured over time via serial dilutions. The black line represents the limit of detection for the first serial dilution. C-D. Shannon diversity (C) and richness (D) in stool communities over time. B-D. Each symbol represents a stool sample from an individual mouse with the lines representing the median value for each treatment group. The transparancy of the symbol corresponds to the day post-infection. Asterisks indicate time points with significant differences between groups by a Kruskall-Wallis test with a Benjamini-Hochberg correction for testing multiple times points. Colored rectangles indicates the 1-day PEG treatment period for applicable groups. The data presented are from a total of `r post_CDI_total_exp` separate experiments.
\newpage

\includegraphics{figure_6.pdf}
**Figure 6. For 1-day PEG treatment post *C. difficile* challenge mice that also receive an FMT only some bacterial genera were restored.** A. PCoA of Bray-Curtis distances from stool samples collected over time as well as the FMT solution that was administered to one treatment group. Each circle represents an individual sample, the transparency of the circle corresponds to day post-infection as shown in Fig. 6C-D. B. Median relative abundances of 2 genera that were significantly different over multiple time points in mice that were administered either FMT or PBS solution via gavage C. Median relative abundances of the top 6 out of `r post_group_multi_day` genera that were significant over multiple timepoints, plotted over time (Data Set S1, Sheet X). For B-C, colored rectangles indicates the 1-day PEG treatment period for applicable groups. Gray horizontal lines represent the limit of detection.

@@

\includegraphics{figure_7.pdf}
**Figure 7. Specific microbiota features associated with prolonged *C. difficile* colonization in PEG treated mice.** A. Top ten bacteria that contributed to the random forest model trained on five day post-infection community relative abundance data to predict wehther mice would still be colonized with *C. difficile* 10 days post-infection. The median (point) and interquantile range (lines) change in AUROC when the bacteria is left out of the model is shown. B. The median relative abundances at 5 days post-infection of the top ten bacteria that contributed to the random forest classification model. Color indicates whether the mice were still colonized with *C. difficile* 10 days post-infection and the black horizontal line represents the median relative abundance. Each symbol represents a stool sample from an individual mouse and the shape of the symbol indicates whether the PEG-treated mice recieved a 5-day (Fig. 1-3), 1-day (Fig. 4) or post-CDI PEG (Fig. 5-6) treatment. C. The median relative abundances of the 5 genera with greater than 1% median relative abundance in the stool community over time. For B-C, the gray horizontal line represents the limit of detection.
\newpage

\includegraphics{figure_8.pdf}
**Figure 8. Schematic summarizing findings.** A. The gut microbiota of our C57Bl/6 mice is restant to *C. difficle* but treatment with either the antibiotic, clindamycin, or the osmotic laxative, PEG 3350 renders the mice susceptible to *C. difficile* colonization. Recovery of colonization resistance in clindamycin-treated mice is relatively straightforward and the mice clear *C.difficile* within 10 days post-infection. However, for mice that recieved a 5-day PEG treatment or a 1-day PEG treatment after *C. difficile* challenge recovery of colonization resistance is more uncertain because mice were still colonized with *C. difficile* 30 days post-infection in the case of several PEG treatments. We found increased *Porphyromonadaceae* and *Lachnospiraceae* were associated with recovery of colonization resistance, while increased *Enterobacteriaceae* and *Bacteroides* were associated with prolonged *C. difficle* colonization. B.Considering that most hospitals recommend not performing *C. difficile* testing on patients taking laxatives and laxatives are used when administering fecal microbiota transplants via colonoscopy to patients with recurrent CDIs. further studies are needed to evaluate if laxatives impact human microbiota colonization resistance. Further studies are needed to understand the impact of osmotic laxatives on *C. difficile* colonization resistance and clearance in human patients.
\newpage

\includegraphics{figure_S1.pdf}
**Figure S1. Microbiota dynamics post-infection in the 5-day PEG treatment plus 10-day recovery mice.** A. *C. difficile* CFU/g over time in the stool samples collected from 5-day PEG treatment plus 10-day recovery mice. Same data presented in Fig. 1C, but the data for the other 3 treatment groups have been removed. B. Median relative abundances of 8 bacterial genera from day 0 post-infection onward from the 10-day recovery mice. We analyzed samples from day 0 and day 8 post-infection, which represented the the time points where mice were challenged with *C. difficile* and when the median relative *C. difficile* CFU stabilized for the group using the paired Wilcoxan signed-rank test, but no genera were signifcant after Benjamini-Hochberg correction.
\newpage

\includegraphics{figure_S2.pdf}
**Figure S2. PEG treatment still has a large impact on the mucosal microbiota 6 days post-infection** A. The   
\newpage

@@

\includegraphics{figure_S4.pdf}
**Figure S4. Specific OTUs associated with clearance by 10 days post-infection that are mostly absent in mice with prolonged *C. difficile* colonization. Ex. *Muribaculum intestinale*.** A.
\newpage
